# Data Access

```{r}

```

## The Data Warehouse
The 'data warehouse' is actually a collection of disparate databases hosted by IDEA Public Schools.  Some data bases are used to host the original source data, some are used as the data model for production applications

```{r display_dw_tables, echo=FALSE, warning=FALSE}

rna_dw_usage <- readxl::read_excel("DataWarehouse Usage.xlsx") %>% 
  janitor::clean_names() %>% 
  select(r_and_a_uses = r_a_uses_x,
         server_name, 
         database_name, 
         table_name, 
         column_name, 
         data_type, 
         constraint) %>% 
  mutate(
         across(where(is.character), as_factor),
         r_and_a_uses = !is.na(r_and_a_uses))


rna_dw_usage %>% 
  janitor::clean_names("title") %>% 
  DT::datatable(filter = "top", rownames = FALSE)
```


### IP Addresses of current servers

For the fastest performance when working remotely it is best to directly access a database by using the IP address of the server rather than than the URL, which often has multiple hops over the network when resolving the DSN. Here's a list of known IP Addresses and server names.

| IP Adress   | Server Name       |
|-------------|-------------------|
|172.24.16.195|   795184-LOGI     |
|172.24.16.208|   795198-CUSTAPPS |
|172.24.16.209|   795199-CUSTAPSV |
|172.24.16.166|   887192-SQLDS    |
|172.24.16.193|   795182-SQLCUSAP |
|172.24.16.125|   795178-DISERVIC |
|172.24.16.207|   795197-DIONLINE |
|172.24.16.198|   964592-SQLDS    |
|172.24.17.56 |   1123130-SDPDTMP |
|172.24.16.83 |   1064618-SQLDI   |
|172.24.16.216|   1054148-FINAPP  |
|172.24.17.111|   1064596-SQLRPT2 |
|172.24.17.110|   1064599-SQLRPTH |
|172.24.16.80 |   1064611-SRC4    |
|172.24.16.206|   795196-HQSTAGIN |
|172.24.16.196|   795186-HQPROD01 |
|172.24.16.197|   795187-HQSQA01  |
|172.24.16.201|   795191-AUTOMATI |
|172.24.16.82 |   1064615-API1    |
|172.24.16.110|   1017144-SQLRPT  |
|172.24.16.205|   795195-HQSQLSA1 | 
|172.24.16.81 |   1064613-SQLPRD2 |
|10.255.88.201|   HQVSQLDS        |
|172.24.17.160|   1065574-SQLPRD1 |
|172.24.16.210|   1044407-SQLDS2  |
|10.255.88.176|   HQVSDBACKUPS    |
|172.24.16.79 |   924794-AUXAUTO  |
|10.255.88.152|   SQLBI           |

: List of Windows Servers hosting SQLServer DBs and their IP Addresses
