# Data Access

```{r}

```

## The Data Warehouse
The 'data warehouse' is actually a collection of disparate databases hosted by IDEA Public Schools.  Some data bases are used to host the original source data, some are used as the data model for production applications

```{r display_dw_tables, echo=FALSE, warning=FALSE}

rna_dw_usage <- readxl::read_excel("DataWarehouse Usage.xlsx") %>% 
  janitor::clean_names() %>% 
  select(r_and_a_uses = r_a_uses_x,
         server_name, 
         database_name, 
         table_name, 
         column_name, 
         data_type, 
         constraint) %>% 
  mutate(
         across(where(is.character), as_factor),
         r_and_a_uses = !is.na(r_and_a_uses)) %>% 
  distinct()


rna_dw_usage %>% 
  janitor::clean_names("title") %>% 
  DT::datatable(filter = "top", rownames = FALSE, )
```


### IP Addresses of current servers

For the fastest performance when working remotely it is best to directly access a database by using the IP address of the server rather than than the URL, which often has multiple hops over the network when resolving the DSN. Here's a list of known IP Addresses and server names.

| IP Adress   | Server Name       |
|-------------|-------------------|
|[REDACTED-IP]|   [REDACTED-HOST]     |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST]    |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST]    |
|[REDACTED-IP] |   [REDACTED-HOST] |
|[REDACTED-IP] |   [REDACTED-HOST]   |
|[REDACTED-IP]|   [REDACTED-HOST]  |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP] |   [REDACTED-HOST]    |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST]  |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP] |   [REDACTED-HOST]    |
|[REDACTED-IP]|   [REDACTED-HOST]  |
|[REDACTED-IP]|   [REDACTED-HOST] | 
|[REDACTED-IP] |   [REDACTED-HOST] |
|[REDACTED-IP]|   [REDACTED-HOST]        |
|[REDACTED-IP]|   [REDACTED-HOST] |
|[REDACTED-IP]|   1044407-SQLDS2  |
|[REDACTED-IP]|   [REDACTED-HOST]    |
|[REDACTED-IP] |   [REDACTED-HOST]  |
|[REDACTED-IP]|   [REDACTED-HOST]           |

: List of Windows Servers hosting SQL Server Databases and their IP Addresses


### Databases that R&A frquently uses. 
```{r}
rna_dw_usage %>% filter(r_and_a_uses) %>% 
  select(server_name, database_name) %>% 
  distinct() %>% 
  janitor::clean_names("title")
```
: List of Windows Servers hosting SQLServer DBs and their IP Addresses

