# Measures

## The Students Table

The Students table is probably the most important table of all. It contains historical student personal data since IDEA's inception. 

We use this table to know to what subpopulation a student belongs to. Also, we can link this table to other student data using the _[StudentNumber]_ field.

**Table:** [[REDACTED-HOST]].[PROD1].[Schools].[Students]

**Main fields:** [AcademicYear] [StudentNumber] | [SchoolNumber] | [GradeLevelID] | [EnrollmentStatus] | [Gender] | [SPED] | [ELLCode] | [PrimaryDisabilityCode ] | [SecondaryDisabilityCode ] | [TertiaryDisabilityCode ] | [EconomicDisadvantageCode] | [FederalHispanicFlag] | [FederaRaceI] | [FederaRaceA] | [FederaRaceB] | [FederaRaceP] |  [FederaRaceW] |[EntryDate] | [ExitDate] | [RowIsCurrent]

**Who is what?:** 

* **_[EnrollmentStatus]:_** 
  + If EnrollmentStatus = 0, it means that the student is/was an active student for the corresponding academic year,
  + If EnrollmentStatus = 2, it means the student left the district before completing the academic year,
  + If EnrollmentStatus = 3, it means the student graduated from IDEA's high school.

* **_[ELLCode]:_** English Language Learner (ELL) A.K.A. Limited English Proficiency (LEP)
  + If ELLCode = 0, student **is not** identified as an ELL student,
  + If ELLCode = 1, the student **is** identified as an ELL student,
  + If ELLCode not in (0,1), it means the student is not currently identified as an ELL but was identified before as one.

* **_[SPED]:_** SPED students are composed of RISE and Life Skills students, not including Critical Student Intervention (CSI) students.
  + If SPED = 0, it means the student is not identified as a SPED student, otherwise then SPED = 1.
  
* **_[InstructionalSettingCode]:_** Students who are classified as SPED may receive additional services. Their services can be found in the **InstructionalSettingDescription** column, which is linked to a specific [PEIMS code](http://ritter.tea.state.tx.us/peims/standards/1314/c035.html):

  InstructionalSettingCode | Description
  -------------------------|-------------
 0 | No additional SPED services
 1 | Homebound
 40 | Mainstream
 41 | Resource Room/Services < 21%
 42 | Resource Room/Services 21- <50%
 43 | Resource Room/Services 50- 60%
 44 | Resource Room/Services > 60%
 
  + If InstructionalSettingCode = 43 or 44, the student is typically in RISE.

* **_[DisabilityCode]_**

  DisabilityCode | Description
  ---------------|-------------
 0 | No disability
 1 | Orthopedic impairment
 2 | Other health impairment
 3 | Auditory impairment
 4 | Visual impairment
 5 | Deaf-Blind
 6 | Intellectual Disability
 7 | Emotional disturbance
 8 | Learning disability
 9 | Speech impairment
 10 | Autism
 13 | Traumatic brain injury
 14 | Noncategorical early childhood

```{r disability, echo=TRUE, eval=FALSE}
SELECT DISTINCT 
 [StudentNumber], 
 [PrimaryDisabilityCode] AS [DisabilityCode], 
 [PrimaryDisabilityDescription] AS [DisabilityDescription]

 FROM (

SELECT DISTINCT [StudentNumber],[PrimaryDisabilityCode], [PrimaryDisabilityDescription]

 FROM [PROD1].[Schools].[Students]
 WHERE AcademicYear='2018-2019'

 UNION ALL

SELECT DISTINCT  [StudentNumber], [SecondaryDisabilityCode], [SecondaryDisabilityDescription]

 FROM [PROD1].[Schools].[Students]
 WHERE AcademicYear='2018-2019'

 UNION ALL

SELECT DISTINCT  [StudentNumber], [TertiaryDisabilityCode], [TertiaryDisabilityDescription]

 FROM [PROD1].[Schools].[Students]
 WHERE AcademicYear='2018-2019') A WHERE [PrimaryDisabilityDescription] <> '' ORDER BY StudentNumber


```

* **_[EconomicDisadvantageCode]:_** A.K.A ECD 
  + If EconomicDisadvantageCode = 0, then the student is _not identified as economically disadvantaged_, otherwise the student is identified as **economically disadvantaged**,
  + If EconomicDisadvantageCode = 1, then the student is identified as an economically disadvantaged and eligible for free meals,
  + If EconomicDisadvantageCode = 2, the student is also identified as an economically disadvantaged but only eligible for reduced-price meals,
  + If EconomicDisadvantageCode = 99, it means the student is identified as an economically disadvantaged one but has another economic disadvantage.


* **_Race / Ethnicity:_** Same like in TX . 

* **_[RowIsCurrent]:_** 

**Context:**

* **Who Counts?** 



* **Who is a 4year 2year?**



* **Persistence Rate by:**

  + **District:** use all data
  +	**Region:** use the [RegionDescription] field
  +	**School:** [SchoolShortName] + [SchoolType]
  +	**Grade Level:** [GradeLevelID]
  +	**Student Type (new or returning):** use [NewStudent] = 1 to identify new students, if zero then the student is considered as a returning student.


* **Other Useful Fields:**
  + Use [PersistenceCode] or [PersistenceCategory] to know student and family reasons on why a student left the district.
  + [LeaverWeek] to know the week when the student left the district.

* **SQL Code Example:**

```{r example1, echo=TRUE, eval=FALSE}
SELECT 1 -  AVG(CAST([FDOPLEAVER] AS NUMERIC))
FROM [`[REDACTED-HOST]`].[Persistence].[dbo].[PersistenceCode]
WHERE [FDOPCOUNTP] = 1
```

Within this chapter we will find the measures IDEA Public Schools stakeholders talk are the most interested about.

## Student Persistence

Student Persistence is one of the most important measures. Chiefs, VPs, and School Leaders are always monitoring this metric which gives us a good idea of our efforts to provide high quality service to our students and their families. We say a student persisted if the student was enrolled and attended the school on the “First Day Of Persistence (FDOP)” for two consecutive years. 

Example: John enrolled IDEA Public Schools on August 19, 2019 and attended to the school the FDOP which is the Monday after the official first week of school. John completes the 2019-2020 academic year and by August 17, 2020 (FDOP) John enrolls again and attends to the school. If John was enrolled but did not attend on the FDOP and he decides to come back until September, then John won’t count as a student who persisted.

**Database:** [[REDACTED-HOST]].[Persistence]

**Main metrics:** District, Regional, School, Subpopulation, Grade Level

**General Formula:** 1 – (Leavers/All Students)

**Tables:**

* **_[dbo].[PersistenceCode]:_** This table is used to calculate persistence rate for the current academic year. It shows data for the most recent week. We will see different week numbers, but this is because the first day of school is different across the district.

* **_[dbo].[PersistenceCodeHistorical]:_** This table shows the final persistence data (as of last week of persistence). Here we will find historical data starting from 2019-2020. The table will be updated every year.

* **_[dbo].[PersistenceHistorical]:_** As the previous table, this one contains final persistence data. The table will not be updated but it includes 2018-2019 data.

**Context:**

* **Who Counts?** 

To calculate persistence, we need to first filter [FDOPCOUNTP] = 1 (FDOP First Day Of Persistence). The number of students resulting from this operation would be our denominator.

* **Who is a Leaver?**

If FDOPLEAVER equals 1, then that student is considered as a leaver otherwise is a current student. All summer leavers (June – August) from [SummerLeaver] = 1, are marked as FDOPLEAVER = 1. All leavers should have an [EnrollmentStatus] =2. The number of students resulting from this operation would be our numerator.

* **Persistence Rate by:**

  + **District:** use all data
  +	**Region:** use the [RegionDescription] field
  +	**School:** [SchoolShortName] + [SchoolType]
  +	**Grade Level:** [GradeLevelID]
  +	**Student Type (new or returning):** use [NewStudent] = 1 to identify new students, if zero then the student is considered as a returning student.


* **Other Useful Fields:**
  + Use [PersistenceCode] or [PersistenceCategory] to know student and family reasons on why a student left the district.
  + [LeaverWeek] to know the week when the student left the district.

* **SQL Code Example:**

```{r example2, echo=TRUE, eval=FALSE}
SELECT 1 -  AVG(CAST([FDOPLEAVER] AS NUMERIC))
FROM [`[REDACTED-HOST]`].[Persistence].[dbo].[PersistenceCode]
WHERE [FDOPCOUNTP] = 1
```

## College Application & Matriculation Metrics

To close the opportunity gap, IDEA Public Schools is committed to a vision of College for All Children. The College Success Team (CST) lead the initiative of monitoring and identifying the best College/University for each senior student, this is possible with data dashboards that use [Naviance](https://www.hobsons.com/solution/naviance/) information.

**Database:** [[REDACTED-HOST]].[PROD1].[Colleges]

**Main metrics:** Percent at least 1 application, Percent at least 1 submission, Percent at least 1 acceptance

**Most important columns:** [Stage] | [ResultCode]

**Tables:**

* **_[Colleges]:_** Here we will find all colleges/universities. What makes this table important is the [CEEB] field which is an ID that can help us connect distinct college related data sources like Naviance with National Student Clearinghouse data!

* **_[CollegeTuition]:_** Besides having in-state tuition and out-of-state tuition data, we can also find two other unique codes ([OPEID] and [ACTCode]), that will aid us to college data across different data sources. 

* **_[EDocs]:_** Details about  application submitted e-documents.

* **_[StudentCollegeApplication]:_** This is the main table. We use it to calculate most of the metrics the College Application & Matriculation dashboard has. We can find data from 2018 up to the current academic year.

* **_[StudentCollegeApplicationsSummary]:_** 

* **_[StudentScholarships]:_** 


**Context:**

* **Who Counts?** 



* **Who is a 4year 2year?**



* **Persistence Rate by:**

  + **District:** use all data
  +	**Region:** use the [RegionDescription] field
  +	**School:** [SchoolShortName] + [SchoolType]
  +	**Grade Level:** [GradeLevelID]
  +	**Student Type (new or returning):** use [NewStudent] = 1 to identify new students, if zero then the student is considered as a returning student.


* **Other Useful Fields:**
  + Use [PersistenceCode] or [PersistenceCategory] to know student and family reasons on why a student left the district.
  + [LeaverWeek] to know the week when the student left the district.

* **SQL Code Example:**

```{r example3, echo=TRUE, eval=FALSE}
SELECT 1 -  AVG(CAST([FDOPLEAVER] AS NUMERIC))
FROM [`[REDACTED-HOST]`].[Persistence].[dbo].[PersistenceCode]
WHERE [FDOPCOUNTP] = 1
```


## Critical Student Intervention (CSI) Identification {#csi}

To address the achievement gap among students, Dolores Gonzalez, our Chief Program Office introduced the Critical Student Intervention (CSI) program in 2014. Currently, along with her team, Tricia Lopez our VP of Special Programs lead IDEA Public Schools in the effort of reducing the performance gaps. To help Tricia's team, the Software Development team put together a table where we can find the students who are enrolled in a CSI math and/or reading intervention program. 

**Database:** [[REDACTED-HOST]].[PROD1].[Schools]

**Main metrics:** Percent of CSI students by school

**Most important columns:** [ProgramID] | [InterventionType]

**Tables:**

* **_[StudentCSI]:_** This table will aid us to identify the students who are/were enrolled in math or reading CSI intervention programs. Since sometimes there is need of data adjustments from one year to another, the [ProgramID] number might change. However, this should not be a problem because we can always use the [InterventionType] field to correctly associate an intervention program to either math or reading.

* **SQL Code Example:**

```{r csi1, echo=TRUE, eval=FALSE}
SELECT DISTINCT AcademicYear, [Subject], ProgramID
FROM (SELECT [AcademicYear]
      ,[ProgramID]
      ,CASE 
	  WHEN [InterventionType] LIKE '%math%' THEN 'Math'
	  WHEN [InterventionType] LIKE '%reading%' THEN 'Reading' ELSE NULL END AS [Subject]
  FROM [[REDACTED-HOST]].[PROD1].[Schools].[StudentCSI]
  WHERE [AcademicYear] = '2018-2019' AND (InterventionType LIKE '%math%' OR InterventionType LIKE '%reading%')) AS A
GROUP BY AcademicYear, [Subject], ProgramID
```

* **_[StudentCSIDetails]:_** Here we find detailed information about the students who are or were part of a CSI program. We can either use this table to get a student count per school, or we could match the **_[StudentCSI]_** table data to the Students table and get the percent of CSI students per school.

* **SQL Code Example:**

```{r csi2, echo=TRUE, eval=FALSE}
SELECT A.StudentNumber, A.SchoolNumber, B.SchoolName, D.[Subject]
FROM [[REDACTED-HOST]].[PROD1].[Schools].[Students] AS A
INNER JOIN [[REDACTED-HOST]].[PROD1].[Schools].[Schools] AS B
ON A.SchoolNumber = B.SchoolNumber
LEFT JOIN [[REDACTED-HOST]].[PROD1].[Schools].[StudentCSI] AS C
ON A.StudentNumber = C.StudentNumber AND A.AcademicYear = C.AcademicYear
LEFT JOIN 
(SELECT DISTINCT AcademicYear, [Subject], ProgramID
FROM (SELECT [AcademicYear]
      ,[ProgramID]
      ,CASE 
	  WHEN [InterventionType] LIKE '%math%' THEN 'Math'
	  WHEN [InterventionType] LIKE '%reading%' THEN 'Reading' ELSE NULL END AS [Subject]
  FROM [[REDACTED-HOST]].[PROD1].[Schools].[StudentCSI]
  WHERE [AcademicYear] = '2020-2021' AND (InterventionType LIKE '%math%' OR InterventionType LIKE '%reading%')) AS A
GROUP BY AcademicYear, [Subject], ProgramID) AS D
ON C.ProgramID = D.ProgramID
WHERE A.AcademicYear = '2020-2021'
```

###Specific Projects


##Teacher Hiring
Analyzing correlations and predictions between pre-selected hiring measures and teacher performance.


#A. Jobvite Data
          *Obtained from HA team. It is the data from the candidates who apply for IDEA jobs through Jobvite. Not all candidates applied through Jobvite in the past, but HA is trying to streamline the application process and require all applicants for all jobs (at all campuses) to apply through a central system, which is Jobvite. A particular Jobvite file is labeled with a year, such as "Jobvite 19-20", and it will have column "Candidate Submit Date", which is a range from October 2019 to July 2019 (October to July, proceeding the year the applicant will likely begin work). An applicant from that range will most likely have a Start Date of August 2019 (or sometime in the calendar year 2019, starting AFTER the 19-20 Academic Year). However, there are cases when applicants will, for example, apply during the 10/2018 - 07/2019 cycle (Jobvite 19-20), but will not have a Start Date until 2020 or 2021 (so they do not start the following FAll after the application cycle). 
          This data has columns including Candidate Full Name, Candidate Email (personal email used for the application), and some of the pre-hire selection measures, such as GPA. However, all 5 of the pre-hire selection measures are not included yet (Stem Major, Teach For America status, Experience prior to IDEA, GPA, and Teacher Certification). The data also has no Employee IDs yet, as this is a list of candidates. There is a Hire Yes/No column, which has "Yes" for those candidates who were offered a position and "No" for those who were not offered a position. Also, there is a Requisition ID which is basically an application ID number, and it is unique per application. If a single candidate applies for more than one position, the candidate will have a unique ID for each application. Below are some bullet points describing Jobvite and Teacher Export (TCP) data.
          
  Jobvite Data: applications for Teacher Roles through Jobvite (when filtering for Category = Teaching)
•	Contains those who are Hired = Yes and Hired = No. 
   o	“Hired = Yes” means they were offered the position, but does not indicate if they actually accepted it or not. 
   o	Each person may apply for more than one job opening, and if they do, they will have a unique Requisition ID for each application (even though it is the same candidate applying for each separate position).
•	Has Personal Email and Full Name
•	Does NOT have IDEA Work Email, nor EmployeeID
•	Contains everyone who applied from October to the following July, for the subsequent August Start Date (or start dates after that year).
•	Most people who applied in Jobvite 19-20 (10/18 to 7/19), started in Fall 2019 (19-20 Academic Year) 
   o	But a few of those started in Fall 2020 and Fall 2021.

1)	19-20 – 2nd year of Jobvite data (18-19 is first and starting year of Jobvite data)
•	(renamed Jobvite 18-19 SY by Brittany; email 10-12-21)
•	Contains Tentative Start Date column – shows which month/year a new hire started working (doesn’t have to be immediately following that year’s hiring cycle)

2)	20-21 – 3rd year of Jobvite data 
•	(renamed Jobvite 18-19 SY by Brittany; email 10-12-21)
•	Contains Tentative Start Date column

3)	21-22 – 4th year of Jobvite data
•	(renamed Jobvite 18-19 SY by Brittany; email 10-12-21)
•	Contains Tentative Start Date column

#B. TCP/Teacher Export Data

TCP/Teacher Export Data: Teacher Performance data.
•	Includes TCP Level Placements, TCP Composite Score, and TCP component scores (all 5 components). 
   o	GET (35%)
      	Manager and Self Guidepost Ratings contain 6 components
•	Lesson Planning & Delivery
•	Culture, Etc.
      	Components 1-5 are counted, and Component 6 (Core Values) is calculated outside of this score (it is 1 of 5 Components of TCP, worth 5% on its own)
      	Each Guidepost Component has 4-5 items listed under each as 1A, 1B, etc.
   o	Parent Survey (5%)
      	Show up as BLANK if <10 surveys were completed per teacher
   o	Student Survey (5%)
      	Shows up as BANK if:
•	 <10 surveys were completed per teacher
   o	Some campuses encourage students and parents to fill out the surveys, and some do not
•	Teacher taught SPED, pre-K, K, 1st, or 2nd Grades (those students are not able to take the survey/or too young).
   o	Student Achievement (50%)
      	Testing such as STAAR, etc.
   o	Core Values (5%)
•	Contains IDEA Work Email and EmployeeID
•	Contains First Name and Last Name (but not Full Name)
•	There is a 2 year HOLD for the pandemic (we don’t know who was “held”), meaning someone at a particular level (E.g., Level 5) could stay at that level for 2 years without having a composite score that is as high as that level. Usually, the hold is for 1 year.
•	If someone got a 4.5+ composite TCP score for 2 consecutive years in a row, they can move to a Level 5 TCP without having the required years of Experience.
•	For those who do not have Student Achievement scores, the highest TCP Level they can reach is a Level 3 (since Student Achievement is 50% of the score).
•	If a person is missing one of the other 4 TCP components, such as Parent Survey, the points that were from that score are reallocated to the other components, weights the in the same ratios as the other components (so it’s not an equal allocation, meaning you don’t divide the percentage 4 equal ways).

1)	18-19  
•	From Alex, see email

2)	19-20  
•	(“Maura Report”, from Alex, new sheet on the Teacher Export 18-19 file from Alex, see email).
•	Need to request this from Blanca (data table we are already using for TSLIP grant, that we can re-use for Teacher Hiring Round 3).
3)	20-21 – 3rd year of Jobvite data 
•	from Blanca, contains ONLY those with “Teacher” roles OR New Hires and Promotions; all co-teachers, etc., have been removed for that academic year 
•	Contains some who started work in 21-22 but are New Hires or Promotions, so rated Level 1 TCP (but didn’t work in the 20-21 school year

4)	21-22 – NO
•	Not available yet! 
•	No one was rated for this school year because it’s not over yet, and the TCP ratings/placements have not yet been conducted. Will likely be finalized in Fall 2022 by early-October.


#Data Matching
Jobvite 19-20 to Teacher Export 19-20
Jobvite 20-21 to Teacher Export 20-21
Jobvite 21-22 to Teacher Export 21-22 – Can’t do this! TE 21-22 NOT available yet!
•	Thus, we will be looking at 2 years of data:
   o	Jobvite 19-20 to Teacher Export 19-20
   o	Jobvite 20-21 to Teacher Export 20-21
      	Removing Jobvite 19-20 Hires who did not have a Start Date of Fall 2019
      	Removing Jobvite 20-21 Hires who did not have a Start Date of Fall 2020
      	Removing all Teacher Export 19-20 and Teacher Export 20-21, who did not have a Job Title of Teacher
      	Removing all Teacher Export 19-20 and Teacher Export 20-21, who had Internal Role as:
•	SPED CP
•	SPED AC
•	PE CP
•	PE CP
•	(This is due to SPED and PE having state mandated license requirements that are not required for the other subjects, so these should be analyzed separately).
      	Removing Teacher Export 19-20 Teachers who did not work in 19-20 Academic Year 
•	(Removing New Hires, Promotions starting in Fall 2020, TCP Level 1)
      	Removing Teacher Export 20-21 Teachers who did not work in 20-21 Academic Year
•	(Removing New Hires, Promotions starting in Fall 2020, TCP Level 1)

#Join Tables
How to join? 
•	Jobvite has no Employee ID & no Work Email and Teacher Export has Employee ID and Work Email.
•	Possibly join through PROD1.Staffing.Employees table, which contains First Name, Last Name, Birthday, Employee ID, and Work Email
        
          


##Camp Rio

##TSLIP

Grant that contains 4 key components/areas for research questions. Main points of analysis: Level 5 Pilot Program and "Best Practices" library; TCP Level percentages by District/Region/Campus/School and look at percentage and distribution of TCP Level 5 teachers across schools and possibly down to grade level.
