# Visualization Standards

## IDEA branding

When producing a visualization for presentation or publication, use the [Brand Guideline](https://ideapublicschoolsorg.sharepoint.com/sites/MCE/SitePages/Marketing-Team.aspx) standards established by the [Marketing team](https://ideapublicschoolsorg.sharepoint.com/sites/MCE/SitePages/Marketing-Team.aspx). Many of these guidelines, particularly colors, have already been incorporated into an R package, `ideacolors`.


### The `ideacolors` package

The `ideacolors` package is used in conjunction with `ggplot2` to produce graphs and visualizations consistent with the primary and secondary brand color schemes. For more information, consult the `ideacolors` [documentation](https://idea-analytics.github.io/ideacolors/).

To install the package, use the following code:

```{r ideacolors, echo=TRUE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("idea-analytics/ideacolors")
```

## Race/Ethnicity Colors

When creating visuals of Race/Ethnicity we want to be consistent with the colors chosen for each race/ethnicity. Additionally, we want to ensure that the colors are not offensive to any group. Below are the colors that were used for all of the Diversity, Equity, and Inclusion (DEI) work that has been completed in the last couple of months of 2022.

  + **American Indian/Alaskan Native:** idea_colors$melon (#F9A054)
  + **Asian:** idea_colors$cyan (#53B4CC)
  + **Black:** idea_colors$magenta (#EE3A80)
  + **Hispanic:** idea_colors$blue (#0079C1)
  + **Multi-racial:** idea_colors$lime (#A4C364)
  + **Other:** idea_colors$yellow (#FFDE75)
  + **Pacific Islander:** idea_colors$darkblue (#1A4789)
  + **White:** idea_colors$coolgray (#626363)

### Typography

The primary brand font is Proxima Nova; the secondary fonts are Arial or Helvetica. Before using the code, ensure that these fonts have already been installed. Then, to add these fonts to a visualization, use the following code:

```{r fonts, echo=TRUE, eval=FALSE}
# install.packages("showtext")
library(showtext)

showtext_auto()
font_add("proxima",
         regular = "proximanova-regular.ttf",
         bold = "proximanova-extrabold.ttf",
         italic = "proximanova-regularit.ttf",
         bolditalic = "proximanova-extraboldit.ttf")
```

This code will apply the font consistently to all text in any visualization created for that R session.

