# Visualization Standards

## IDEA branding

When producing a visualization for presentation or publication, use the [Brand Guideline](https://ideapublicschoolsorg.sharepoint.com/sites/MCE/SitePages/Marketing-Team.aspx) standards established by the [Marketing team](https://ideapublicschoolsorg.sharepoint.com/sites/MCE/SitePages/Marketing-Team.aspx). Many of these guidelines, particularly colors, have already been incorporated into an R package, `ideacolors`.


### The `ideacolors` package

The `ideacolors` package is used in conjunction with `ggplot2` to produce graphs and visualizations consistent with the primary and secondary brand color schemes. For more information, consult the `ideacolors` [documentation](https://idea-analytics.github.io/ideacolors/).

To install the package, use the following code:

```{r ideacolors, echo=TRUE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("idea-analytics/ideacolors")
```

### Race/Ethnicity Colors

When creating visuals of Race/Ethnicity we want to be consistent with the colors chosen for each race/ethnicity. Additionally, we want to ensure that the colors are not offensive to any group. Below are the colors that were used for all of the Diversity, Equity, and Inclusion (DEI) work that has been completed in the last couple of months of 2022.

  + **American Indian/Alaskan Native:** idea_colors$melon (#F9A054)
  + **Asian:** idea_colors$cyan (#53B4CC)
  + **Black:** idea_colors$magenta (#EE3A80)
  + **Hispanic:** idea_colors$blue (#0079C1)
  + **Multi-racial:** idea_colors$lime (#A4C364)
  + **Other:** idea_colors$yellow (#FFDE75)
  + **Pacific Islander:** idea_colors$darkblue (#1A4789)
  + **White:** idea_colors$coolgray (#626363)
  
Suppose your data frame, `df_reading_levels`, contains information about Academy students' grade-level reading equivalents `GradeLevelEquivalent` with race and ethnicity in the `RaceEthnicity` column. To create a visualization of reading levels using the custom palette, you'll need to code the palette in three places:

1. as a vector of colors, outside of the plotting function
2. as an aesthetic, inside the initial plotting call
3. as a value, inside a manual scale function

Let's define the palette as a vector. You may need to adjust the names of the race or ethnicity to match the factors in your data frame.

```{r vector_colors, eval=FALSE}
# step 1
race_ethnicity_legend <- c("American Indian" = idea_colors$melon,
                           "Asian" = idea_colors$cyan,
                           "Black" = idea_colors$magenta,
                           "Hispanic" = idea_colors$blue,
                           "Other" = idea_colors$yellow,
                           "Pacific Islander" = idea_colors$darkblue,
                           "Two or more races" = idea_colors$lime,
                           "White" = idea_colors$coolgray)
```

Then, inside a `geom_*()` layer, assign the appropriate column `RaceEthnicity` to the appropriate aesthetic `aes()`. (Depending on your visualization, the aesthetic might be `fill`, `shape`, etc.).

```{r geom_aes, eval=FALSE}
# step 2
ggplot(df_reading_levels) +
  geom_boxplot(aes(x = GradeLevelEquivalent,
                   color = RaceEthnicity))
```

One additional call is needed to force `ggplot()` to use the palette you specified earlier.

```{r scale_color_manual, eval=FALSE}
# step 2
ggplot(df_reading_levels) +
  geom_boxplot(aes(x = GradeLevelEquivalent,
                   color = RaceEthnicity)) +
  # other layers go here... +
  # step 3
  scale_color_manual(values = race_ethnicity_legend,
                     drop = FALSE) +
  labs(color = "Race / Ethnicity")
```

The `scale_color_manual()` call will accept the vector of colors specified earlier in `values`, and `drop = FALSE` forces the legend to show all possible values and colors of `race_ethnicity_legend`, regardless if they appear in the `RaceEthnicity` column. (If the aesthetic if `fill`, then use `scale_fill_manual(values = race_ethnicity_legend, drop = FALSE)`, and so on).

The default legend title is to use the column name. However, the `labs()` call *relabels* the legend title for the `color` aesthetic to "Race / Ethnicity". (If the aesthetic is `fill`, then use `labs(fill = "Race / Ethnicity")`, and so on).

### Grade Level colors

If your visualization or reporting requires grouping across grade levels, using the IDEA polo color for that grade level is a familiar palette for IDEA schools. For more information about uniforms, refer to the [parent guide](https://ideapublicschools.org/parents/uniform-guides/).

The following are the current polo colors for each grade level:

  Grade Level | Color 
  ------------|-------
 PK-5th, 8th | Blue
 6th, 10th | Red
 7th, 11th | Orange
 9th | Green
 12th | Black
 
Clearly, if all grade levels are present on one plot and cannot be individually labeled (say, a scatterplot), then using one color for multiple grade levels is not advised. However, if color coding a document or table where colors can be repeated, then this palette can provide visual cues to contextualize the report.


### Typography

The primary brand font is Proxima Nova; the secondary fonts are Arial or Helvetica. Before using the code, ensure that these fonts have already been installed. Then, to add these fonts to a visualization, use the following code:

```{r fonts, echo=TRUE, eval=FALSE}
# install.packages("showtext")
library(showtext)

showtext_auto()
font_add("proxima",
         regular = "proximanova-regular.ttf",
         bold = "proximanova-extrabold.ttf",
         italic = "proximanova-regularit.ttf",
         bolditalic = "proximanova-extraboldit.ttf")
```

This code will apply the font consistently to all text in any visualization created for that R session.

