--- 
title: "IDEA R&A Manual"
author: "Chris Haid, Edison Coronado, Mishan Jensen, Maura Carter, Steven Macapagal, Aline Orr, Ilissa Madrigal, Marlena Coco"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
url: https://idea-analytics.github.io/r_and_a_manual/
cover-image: 
description: |
  This is IDEA Public Schools' Research & Analysis Team's ever evolving encyclopedia of resources and best practices.
biblio-style: apalike
csl: chicago-fullnote-bibliography.csl
---

# A little about this manual {#about}

Placeholder


## What goes in this manual?
## How do you contribute to this book?
### Adding "Tips" and other callouts

<!--chapter:end:index.Rmd-->


# Contributors {#who}

Placeholder


## About Research and Analysis
## Who we are
### Edison Coronado
### Chris Haid 
### Mishan Jensen, PhD
### Steven Macapagal
### Maura Carter
### Aline Orr, PhD
### Ilissa Madrigal
### Marlena Coco, PhD
## What we've contributed

<!--chapter:end:02-Who_We_Are.Rmd-->


# The Data Warehouse and how to access it

Placeholder


## The Data Warehouse
### Core data warehouse server configuration
#### SRC Servers
#### PROD Servers
#### RS and Hisotrical data Servers
### Server, Table, and Field Lookup
## Accessing the data warehouse from R
### Installation
### Example
#### What if I am pulling down lots of data (say, millions of rows)?
### Finding things in the data warehouse
### Where to learn more

<!--chapter:end:03-Data_Warehouse.Rmd-->


# Key Tables, Measures, and Metrics

Placeholder


## The Students Table
### Common queries
## Schools and regions
### Schools table
### Regions table
## Student Attendance
### Common queries
### Other uses
## Student Persistence
### Student Persistence using Attendance Tables (A.K.A. Cohort Persistence)
## College Application & Matriculation Metrics
## Lottery Data
## Critical Student Intervention (CSI) Identification {#csi}
## Assessment Data
### State of Texas Assessments of Academic Readiness (STAAR)
### District assessments
#### Illuminate data
##### Naming conventions
##### IABWA
##### IABWAItems
#### Other sources
### DIBELS
### Renaissance STAR (also RenStar, Little Star)
#### Where to get RenStar data?
### AP
### IB
#### Coursework and scoring
#### Data
### ACT
### LEAP
#### Student level LEAP
## Student Course Grades
## Teacher, Staff, and Hiring Data
### Compass data
#### Key Tables
### Jobvite Data
### TCP/Teacher Export Data
#### Data Matching
#### Join Tables
## Evaluation Projects
### 21st Century Afterschool Programs
### Camp Rio
### TSLIP
### Charter School Program Grants {#csi_tables}
#### APR Report for June 2022 
##### Promotions
###### Teacher Promotions
###### Teacher-Leader Promotions
#### TCP Retention and TCP Levels
##### Missing TCP Placements
##### Leaver Rates
##### Exit Surveys for Teacher Leavers
#### TCP Levels and Composite Score Correlations

<!--chapter:end:04-Tables-Measures-and-Metrics.Rmd-->


# Projects

Placeholder


## Version control with Github {#section-github}
### Getting Started with Git, Github, and RStudio
#### First steps {#first-steps}
### Feature Branch Worklow
### Example worflow with this manual, or getting your feet wet
#### Get the [R_and_A\_Manual](https://github.com/idea-analytics/r_and_a_manual/) repository URL
#### Clone the repo {.tabset}
##### Comand line
##### RStudio
#### Checkout a branch
##### Command line
##### RStudio
#### Making changes and saving them
#### Command line
#### Maintaining large files
##### Large file storage (LFS)
##### Using `.gitignore`
#### RStudio
#### Merging changes.
## Project Process
### Wrike
#### Onboarding and setup
#### Goal setting and deadlines
#### Defining tasks and dependencies
#### Documenting and evaluating progress towards goal
### GRPI and RASI
### Project operating mechanisms
## ProjectTemplate (for analyses) {#section-ProjectTemplate}
### Installing ProjectTemplate
### Creating a project with ProjectTemplate
### What goes where
## The `renv` package: Ensuring reproducibility
### Setting up `renv` with a new project
### Setting up `renv` with an existing project
### Using `renv`
### Collaborating with `renv`
### `renv` and RStudio Connect
## Publishing Projects
### RStudio Connect
### Power BI
### Team Site (TheHub)
#### Update Team Site Annual Evaluation Plans
## Specific Projects
### Teacher Hiring and TCP
#### Jobvite Data
#### Teacher Export Data
#### TCP Handbook
#### Teacher Interviews/Teacher Manager Interviews
### First Year Teacher Project
#### 3 Years of Data Used in Analysis
#### Project Transfered to Ann Heller
### Math Curriculum Redesign
### TSLIP
### Charter School Program (CSP) Grants

<!--chapter:end:05-Projects.Rmd-->

```{r include=FALSE, cache=FALSE}
# example R options set globally
options(width = 60)

# example chunk options set globally
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
  )
```
# Coding Standards

We will be writing more here in the coming weeks, months, and years. But for all intents and purposes, we should follow the [Tidyverse Style Guide](https://style.tidyverse.org/index.html), which is was originally derived from Google's R Style guide, but in a an amazing case of an [Ouroboros](https://en.wikipedia.org/wiki/Ouroboros), is now informed by the Tidyverse Style Guide.

[![ouroboros: A Snake Eating it's Tail](https://media.istockphoto.com/vectors/uroboros-snake-coiled-in-a-ring-biting-its-tail-scratch-board-black-vector-id1216824086?k=20&m=1216824086&s=612x612&w=0&h=C4GD_7_z3Ma8v3Z9GRSPEkHlbx9cLQ0R7IkBdOl9RuM= "A Snake Eating it's Tail")](https://en.wikipedia.org/wiki/Ouroboros)

## File naming standards. 

We will also come up with more here soon, but in the meantime, Jenny Bryan's guidance is pretty excellent.  [You can read more here](http://www2.stat.duke.edu/~rcs46/lectures_2015/01-markdown-git/slides/naming-slides/naming-slides.pdf)

<!--chapter:end:06-Code-Standards.Rmd-->


# Visualization Standards

Placeholder


## IDEA branding
### The `ideacolors` package
### Race/Ethnicity Colors
### Typography

<!--chapter:end:07-Visual-Standards.Rmd-->


# Models and Methods

Placeholder


## Research design
### Causal inference
#### Potential outcomes
#### Matching models
#### Regression discontinuity
#### Instrumental variables
#### Difference-in-differences
### Experimental design
#### Differences among distributions of groups
#### Differences among levels of variables
## Descriptive statistics
### Numerical summaries - broad principles
### Graphical summaries - broad principles
#### Displaying variation
## Inferential statistics
### Interval estimates
### p-values and hypothesis testing
## Modeling and predictive statistics
### Linear modeling
### Generalized linear modeling
### Machine learning
## Geospatial Methods, or A Gentle Introduction to GIS in R with Simple Features
### What is a simple feature?
#### How are simple features implemented in R
### Let's look at TACO schools
#### Geocoding
##### Coordinate Reference System
#### Plotting
#### Getting drive times
#### Some geographic filtering
### Census Data
###### Getting Basic Data
###### Searching for variables:
### Chloropleth
### Dot-density
## Qualitative Tools
### MaxQDA

<!--chapter:end:08-Models-and-Methods.Rmd-->


# Tips and Tricks

Placeholder


## Data processing 
### De-duplicating rows in a table. 
### Hierarchical aggregation, or the `list-group_by` trick
#### Saving grouping info
### Creating "Scaffolds" when joining data.  
## Statistical pitfalls
### Simpson's Paradox
### Bias-variance tradeoff
## Troubleshooting messages and errors
### Connection to the dashboard

<!--chapter:end:09-Tips_and_Tricks.Rmd-->

```{r include=FALSE, cache=FALSE}
# example R options set globally
options(width = 60)

# example chunk options set globally
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
  )
```
`r if (knitr:::is_html_output()) '
# References {-}
'`

<!--chapter:end:10-References.Rmd-->


# Specific Projects

Placeholder


## We could move content
## File Storage
### R&A Private Group
## Evaluations
### Evaluation Plans 
### 21st Century Afterschool Program
### Camp Rio
### Charter School Program (CSP) Grants
### Teacher and School Leader Incentive Program (TSLIP)
## Other Specific Projects
### Teacher Hiring and TCP
#### Jobvite Data
#### Teacher Export Data
#### TCP Handbook
#### Teacher Interviews/Teacher Manager Interviews

<!--chapter:end:6-Specific-Projects.RMD-->

