--- 
title: "IDEA R&A Manual"
author: "Chris Haid, Edison Coronado, Mishan Jensen, Maura Carter, Steven Macapagal, Aline Orr, Ilissa Madrigal, Marlena Coco"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
url: https://idea-analytics.github.io/r_and_a_manual/
cover-image: 
description: |
  This is IDEA Public Schools' Research & Analysis Team's ever evolving encyclopedia of resources and best practices.
biblio-style: apalike
csl: chicago-fullnote-bibliography.csl
---

# A little about this manual {#about}

Placeholder


## What goes in this manual?
## How do you contribute to this book?
### Adding "Tips" and other callouts

<!--chapter:end:index.Rmd-->


# Who we are {#who}

Placeholder


## About Research and Analysis
### Edison Coronado
### Chris Haid 
### Mishan Jensen, PhD
### Steven Macapagal
### Maura Carter
### Aline Orr, PhD
### Ilissa Madrigal
### Marlena Coco, PhD

<!--chapter:end:02-Who_We_Are.Rmd-->


# The Data Warehouse and how to access it

Placeholder


## The Data Warehouse
### Core data warehouse server configuration
#### SRC Servers
#### PROD Servers
#### RS and Hisotrical data Servers
### Server, Table, and Field Lookup
## Accessing the data warehouse from R
### Installation
### Example
#### What if I am pulling down lots of data (say, millions of rows)?
### Finding things in the data warehouse
### Where to learn more

<!--chapter:end:03-Data_Warehouse.Rmd-->


# Key Tables, Measures, and Metrics

Placeholder


## The Students Table
### Common queries
## Schools and regions
### Schools table
### Regions table
## Student Attendance
### Common queries
### Other uses
## Student Persistence
### Student Persistence using Attendance Tables (A.K.A. Cohort Persistence)
## College Application & Matriculation Metrics
## Lottery Data
## Critical Student Intervention (CSI) Identification {#csi}
## Assessment Data
### State of Texas Assessments of Academic Readiness (STAAR)
### District assessments
#### Illuminate data
##### Naming conventions
##### IABWA
##### IABWAItems
#### Other sources
### DIBELS
### Renaissance STAR (also RenStar, Little Star)
#### Where to get RenStar data?
### IB
#### Coursework and scoring
#### Data
## Student Course Grades
## Teacher, Staff, and Hiring Data
### Jobvite Data
### TCP/Teacher Export Data
#### Data Matching
#### Join Tables
## Evaluation Projects
### 21st Century Afterschool Programs
### Camp Rio
### TSLIP
#### APR Report for June 2022 
##### Promotions
###### Teacher Promotions
###### Teacher-Leader Promotions
#### TCP Retention and TCP Levels
##### Missing TCP Placements
##### Leaver Rates
##### Exit Surveys for Teacher Leavers
#### TCP Levels and Composite Score Correlations

<!--chapter:end:04-Tables-Measures-and-Metrics.Rmd-->


# Projects

Placeholder


## Version control with Github {#section-github}
### Getting Started with Git, Github, and RStudio
#### First steps {#first-steps}
### Feature Branch Worklow
### Example worflow with this manual, or getting your feet wet
#### Get the [R_and_A\_Manual](https://github.com/idea-analytics/r_and_a_manual/) repository URL
#### Clone the repo {.tabset}
##### Comand line
##### RStudio
#### Checkout a branch
##### Command line
##### RStudio
#### Making changes and saving them
#### Command line
#### Maintaining large files
##### Large file storage (LFS)
##### Using `.gitignore`
#### RStudio
#### Merging changes.
## Project Process
### Wrike
#### Onboarding and setup
#### Goal setting and deadlines
#### Defining tasks and dependencies
#### Documenting and evaluating progress towards goal
### GRPI and RASI
### Project operating mechanisms
## ProjectTemplate (for analyses) {#section-ProjectTemplate}
### Installing ProjectTemplate
### Creating a project with ProjectTemplate
### What goes where
## The `renv` package: Ensuring reproducibility
### Setting up `renv` with a new project
### Setting up `renv` with an existing project
### Using `renv`
### Collaborating with `renv`
### `renv` and RStudio Connect
## Publishing Projects
### RStudio Connect
### Power BI
### Team Site (TheHub)
## Teacher Hiring and TCP
### Jobvite Data
### Teacher Export Data
### TCP Handbook
### Teacher Interviews/Teacher Manager Interviews
## First Year Teacher Project
### 3 Years of Data Used in Analysis
### Project Transfered to Ann Heller
## Math Curriculum Redesign

<!--chapter:end:05-Projects.Rmd-->

```{r include=FALSE, cache=FALSE}
# example R options set globally
options(width = 60)

# example chunk options set globally
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
  )
```
# Coding Standards

We will be writing more here in the coming weeks, months, and years. But for all intents and purposes, we should follow the [Tidyverse Style Guide](https://style.tidyverse.org/index.html), which is was originally derived from Google's R Style guide, but in a an amazing case of an [Ouroboros](https://en.wikipedia.org/wiki/Ouroboros), is now informed by the Tidyverse Style Guide.

[![ouroboros: A Snake Eating it's Tail](https://media.istockphoto.com/vectors/uroboros-snake-coiled-in-a-ring-biting-its-tail-scratch-board-black-vector-id1216824086?k=20&m=1216824086&s=612x612&w=0&h=C4GD_7_z3Ma8v3Z9GRSPEkHlbx9cLQ0R7IkBdOl9RuM= "A Snake Eating it's Tail")](https://en.wikipedia.org/wiki/Ouroboros)

## File naming standards. 

We will also come up with more here soon, but in the meantime, Jenny Bryan's guidance is pretty excellent.  [You can read more here](http://www2.stat.duke.edu/~rcs46/lectures_2015/01-markdown-git/slides/naming-slides/naming-slides.pdf)

<!--chapter:end:06-Code-Standards.Rmd-->

```{r include=FALSE, cache=FALSE}
# example R options set globally
options(width = 60)

# example chunk options set globally
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
  )
```
# Visualization Standards

## IDEA branding

When producing a visualization for presentation or publication, use the [Brand Guideline](https://ideapublicschoolsorg.sharepoint.com/sites/MCE/SitePages/Marketing-Team.aspx) standards established by the [Marketing team](https://ideapublicschoolsorg.sharepoint.com/sites/MCE/SitePages/Marketing-Team.aspx). Many of these guidelines, particularly colors, have already been incorporated into an R package, `ideacolors`.


### The `ideacolors` package

The `ideacolors` package is used in conjunction with `ggplot2` to produce graphs and visualizations consistent with the primary and secondary brand color schemes. For more information, consult the `ideacolors` [documentation](https://idea-analytics.github.io/ideacolors/).

To install the package, use the following code:

```{r ideacolors, echo=TRUE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("idea-analytics/ideacolors")
```

## Race/Ethnicity Colors

When creating visuals of Race/Ethnicity we want to be consistent with the colors chosen for each race/ethnicity. Additionally, we want to ensure that the colors are not offensive to any group. Below are the colors that were used for all of the Diversity, Equity, and Inclusion (DEI) work that has been completed in the last couple of months of 2022.

  + **American Indian/Alaskan Native:** idea_colors$melon (#F9A054)
  + **Asian:** idea_colors$cyan (#53B4CC)
  + **Black:** idea_colors$magenta (#EE3A80)
  + **Hispanic:** idea_colors$blue (#0079C1)
  + **Multi-racial:** idea_colors$lime (#A4C364)
  + **Other:** idea_colors$yellow (#FFDE75)
  + **Pacific Islander:** idea_colors$darkblue (#1A4789)
  + **White:** idea_colors$coolgray (#626363)

### Typography

The primary brand font is Proxima Nova; the secondary fonts are Arial or Helvetica. Before using the code, ensure that these fonts have already been installed. Then, to add these fonts to a visualization, use the following code:

```{r fonts, echo=TRUE, eval=FALSE}
# install.packages("showtext")
library(showtext)

showtext_auto()
font_add("proxima",
         regular = "proximanova-regular.ttf",
         bold = "proximanova-extrabold.ttf",
         italic = "proximanova-regularit.ttf",
         bolditalic = "proximanova-extraboldit.ttf")
```

This code will apply the font consistently to all text in any visualization created for that R session.


<!--chapter:end:07-Visual-Standards.Rmd-->


# Models and Methods 

Placeholder


## Research design
### Causal inference
#### Potential outcomes
#### Matching models 
#### Regression discontinuity
#### Instrumental variables
#### Difference-in-differences
### Experimental design
#### Differences among distributions of groups
#### Differences among levels of variables
## Descriptive statistics
### Numerical summaries - broad principles
### Graphical summaries - broad principles
#### Displaying variation
## Inferential statistics
### Interval estimates
### p-values and hypothesis testing
## Modeling and predictive statistics
### Linear modeling
### Generalized linear modeling
### Machine learning

<!--chapter:end:08-Models.Rmd-->


# Tips and Tricks

Placeholder


## Data processing 
### De-duplicating rows in a table. 
### Hierarchical aggregation, or the `list-group_by` trick
#### Saving grouping info
### Creating "Scaffolds" when joining data.  
## Statistical pitfalls
### Simpson's Paradox
### Bias-variance tradeoff

<!--chapter:end:09-Tips_and_Tricks.Rmd-->

```{r include=FALSE, cache=FALSE}
# example R options set globally
options(width = 60)

# example chunk options set globally
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
  )
```
`r if (knitr:::is_html_output()) '
# References {-}
'`

<!--chapter:end:10-References.Rmd-->

